Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-05-12 articles/formalism/sections/kernel_discussion.tex
Adding paintbrush figure into main text
Editing pass and notation cleanup.

The real and imaginary part of the kernel $\mathcal{M}_G$ are identical irrespective of changes in the galactic latitude and longitude of the central pixel. This is consistent with the Green's function interpretation of these kernels. Note that these functions are not distorted when a part of the domain overlaps with the poles, as can be seen in the first three rows of \fig{fig:vis_kernel}. Both these facts can be associated with the fact that this function does not depend on the Euler angle $\gamma$. From \eq{eq:qu2eb_convolution_explicit} and \eq{eq:eb2qu_convolution_explicit} it is clear that $\mathcal{M}_r$ and $\mathcal{M}_i$ can be interpreted in the following ways,
%\revisit{Recall that the angle $\gamma$ defines the rotation about the final z-axis which co-alligns the planar axes. Since the function $\mathcal{M}$ does not depend of this final orientation, it implies a rotational invariance of the fields constructed by convolving with this kernel.}
%
\begin{subequations}
\beqry
\bmat E= \textrm{Re}(\mathcal{M}_G) \\ B =\textrm{Im}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=\delta(\hat{n} - \hat{n}_q)\\ U=0 \emat ~;~ \bmat E= -\textrm{Im}(\mathcal{M}_G) \\ B = + \textrm{Re}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=0 \\ U=\delta(\hat{n} - \hat{n}_q) \emat \,,\\
\bmat Q= \textrm{Re}(\mathcal{M}^*_B) \\ U = \textrm{Im}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=\delta(\hat{n} - \hat{n}_e)\\ B=0 \emat ~;~ \bmat Q= -\textrm{Im}(\mathcal{M}^*_B) \\ U = \textrm{Re}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=0 \\ B=\delta(\hat{n} - \hat{n}_e) \emat \,.
\eeqry
\end{subequations}
%
The kernels $\mathcal{D}$ \& $\mathcal{I}$ vary significantly as a function of galactic latitude of the central pixel as seen in the last four columns of \fig{fig:vis_kernel}. These kernels show a two fold symmetry in the vicinity of the poles and this arises due to Euler angle $\gamma \approx 0$ here and therefore $e^{i2(\alpha \pm \gamma)} \approx e^{i2\alpha}$. Note that in this region, the azimuthal profile of the real and imaginary part of these kernels  is similar to $\mathcal{M}_r$ and $\mathcal{M}_i$ respectively.  This explains why the imaginary part of the band limited delta function $\mathcal{I}$ contributes just as much as the real part in these regions. On transiting to lower latitudes, $\mathcal{D}$ quickly transitions to having a four fold symmetry while $\mathcal{I}$ transitions to being dominated by the real part and behaves more like the conventional delta function. This transition can be most easily understood in the flat sky limit where $\gamma = -\alpha$ which leads to the resultant 4 fold symmetry seen for $\mathcal{D}$ owing to $e^{i2(\alpha - \gamma)} =e^{i4\alpha}$ and $\mathcal{I}$ being dominated by the real part owing to $e^{-i2(\alpha + \gamma)} =1 + i0$. Since the flat sky approximation has most validity in the proximity of the equator these limiting tendencies of the respective kernels are seen in the bottom row of \fig{fig:vis_kernel} which depict the kernels evaluated at the equator $b=0^{\circ}$. The middle two row depict the kernels evaluated at a latitudes of $b=87^{\circ}~\&~ 80^{\circ}$ and serve to indicate the rate of this transition. These kernels are invariant under changes in longitude of the central pixel, the latitude being held fixed, as one may have expected.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Quantifying the non-locality of E \& B modes} \label{sec:radial_locality}
\subsection{The non-locality of the real space operators} \label{sec:radial_locality}
%
\begin{figure}[t]
\centering
\includegraphics[width=0.8\columnwidth]{beta_kernel.pdf}
\caption{The figure depicts the radial part of the convolution kernels. These radial function have been evaluated with the band limit fixed at $\ell \in [2,24]$. The vertical dashed line marks the approximate Healpix pixel size of a NSIDE=8, which is the lowest resolution that allows access to $\ell_{\rm max}=24$.}
\label{fig:beta_kernel}
\end{figure}
%
In the previous section we presented a quantitative understanding of the  azimuthal dependence of the real space kernels, providing little insight into their radial dependence. The radial part of these kernels determines the non-locality of the respective operators and encodes all their multipole dependencies. For illustration the radial kernels ${_{\mm}f}, {_{\md}f} ~\&~ {_{\mi}f} $ are evaluated using the respective multipole sums given in \eq{eq:rad_ker_queb} and \eq{eq:f2_rad_ker} in the band limit $\ell \in [2,192]$ and the resultant functions are depicted in \fig{fig:beta_kernel}. 
%
\begin{figure}[t]
\centering
\subfigure{\includegraphics[width=0.96\columnwidth]{rad_ker_fn_of_ellmax.pdf}}
\subfigure{\includegraphics[width=0.48\columnwidth]{rad_ker_i_fn_of_ellmax.pdf}}
\subfigure{\includegraphics[width=0.48\columnwidth]{rad_ker_d_fn_of_ellmax.pdf}}
\caption{The top panel depicts the radial function  ${_{\mm}f}(\beta,\ell_{\rm min},\ell_{\rm max})$ while the bottom left and right panels show the radial functions ${}_{\mi}f(\beta,\ell_{\rm min},\ell_{\rm max}) ~\&~ {}_{\md}f(\beta,\ell_{\rm min},\ell_{\rm max})$ respectively, for fixed $\ell_{\rm min}=2$ and varying $\ell_{\rm max}$ as indicated by their legends. All the curves are normalized such that their maxima is set to unity. The horizontal solid black line marks the location where the amplitude of the respective kernels fall below 1\% of its maximum. The thin slanted dashed gray lines indicate a power law fit (by eye) to the envelope of the radial functions. The thick black short vertical dashed lines indicate the transition points as predicted by the empirically derived relation for the non-locality parameter $\beta_0=min(180,180 \frac{22}{\ell_{\rm max}})$.}
%While the envelopes for function ${_{\mm}f}(\beta)~\&~ {}_{\mi}f(\beta)$ are fit well by the power law $\propto \beta^{-1.5}$, the envelope for the function ${}_{\md}f(\beta)$ is seen to have a slightly steeper slope $\propto \beta^{-1.65}$.}
\label{fig:rad_ker_decay}
\end{figure}
%

${_{\mm}f}$ is the radial part of the kernel that translates the Stokes parameters $Q$ \& $U$ to scalars $E$ \& $B$ and vice versa and it has a vanishing contribution from the regions where ($\beta \rightarrow 0$) and from regions ($\beta \rightarrow \pi$) as seen in \fig{fig:beta_kernel}. This can be understood as arising from the fact that the coordinate dependence of the Stokes parameters cannot uniquely integrated out since the Euler angles $\alpha$ \& $\gamma$ are not uniquely defined at the locations $\beta=0,\pi$. This nature of ${_{\mm}f}$ is critical to ensure that the derived fields have the necessary spin properties.

\revisit{${_{\mi}f}$ is the radial part of the band limited delta function $\mi$ and expectedly has its maxima at $\beta=0$ and decays with increasing angular separation.  ${_{\md}f}$ has a vanishing value in the region where $\beta \rightarrow 0$ however it does not vanish at $\beta \rightarrow \pi$ as seen in \fig{fig:beta_kernel}. The main take away from these observations of the nature of the radial part of the kernel is that the real space operators which decompose the Stokes parameters into parts that correspond to E and B modes are necessarily non-local.}\\ \linebreak
\noindent \textit{The band limit dependence:} 
%It is clear from previous discussions that the scalar field $E$ \& $B$ constructed at a location depends on the Stokes field in the surrounding regions. 
We further quantify this non-locality by studying the radial extent of the kernels and its dependence on the maximum multipole accessible for analysis. To carry out this study we evaluate the radial functions for different values of $\ell_{\rm max}$, while keeping the lowest multipole fixed at $\ell_{\rm min}=2$.

>>>>

The real and imaginary part of the kernel $\mathcal{M}_G$ are identical irrespective of changes in the galactic latitude and longitude of the central pixel. This is consistent with the Green's function interpretation of these kernels. Note that these functions are not distorted when a part of the domain overlaps with the poles, as can be seen in the first three rows of \fig{fig:vis_kernel}. Both these facts can be associated with the fact that this function does not depend on the Euler angle $\gamma$. From \eq{eq:qu2eb_convolution_explicit} and \eq{eq:eb2qu_convolution_explicit} it is clear that $\mathcal{M}_r$ and $\mathcal{M}_i$ can be interpreted in the following ways,
%\revisit{Recall that the angle $\gamma$ defines the rotation about the final z-axis which co-alligns the planar axes. Since the function $\mathcal{M}$ does not depend of this final orientation, it implies a rotational invariance of the fields constructed by convolving with this kernel.}
%
\begin{subequations}
\beqry
\bmat E= \textrm{Re}(\mathcal{M}_G) \\ B =\textrm{Im}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=\delta(\hat{n} - \hat{n}_q)\\ U=0 \emat;\qquad
\bmat E= -\textrm{Im}(\mathcal{M}_G) \\ B = + \textrm{Re}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=0 \\ U=\delta(\hat{n} - \hat{n}_q) \emat \,,\\
\bmat Q= \textrm{Re}(\mathcal{M}^*_B) \\ U = \textrm{Im}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=\delta(\hat{n} - \hat{n}_e)\\ B=0 \emat; \qquad
\bmat Q= -\textrm{Im}(\mathcal{M}^*_B) \\ U = \textrm{Re}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=0 \\ B=\delta(\hat{n} - \hat{n}_e) \emat \,.
\eeqry
\end{subequations}
%
The kernels $\mathcal{D}$ \& $\mathcal{I}$ vary significantly as a function of galactic latitude of the central pixel as seen in the last four columns of \fig{fig:vis_kernel}. These kernels show a two fold symmetry in the vicinity of the poles and this arises due to Euler angle $\gamma \approx 0$ here and therefore $e^{i2(\alpha \pm \gamma)} \approx e^{i2\alpha}$. Note that in this region, the azimuthal profile of the real and imaginary part of these kernels  is similar to $\mathcal{M}_r$ and $\mathcal{M}_i$ respectively.  This explains why the imaginary part of the band limited delta function $\mathcal{I}$ contributes just as much as the real part in these regions. On transiting to lower latitudes, $\mathcal{D}$ quickly transitions to having a four fold symmetry while $\mathcal{I}$ transitions to being dominated by the real part and behaves more like the conventional delta function. This transition can be most easily understood in the flat sky limit where $\gamma = -\alpha$ which leads to the resultant 4 fold symmetry seen for $\mathcal{D}$ owing to $e^{i2(\alpha - \gamma)} =e^{i4\alpha}$ and $\mathcal{I}$ being dominated by the real part owing to $e^{-i2(\alpha + \gamma)} =1 + i0$. Since the flat sky approximation has most validity in the proximity of the equator these limiting tendencies of the respective kernels are seen in the bottom row of \fig{fig:vis_kernel} which depict the kernels evaluated at the equator $b=0^{\circ}$. The middle two row depict the kernels evaluated at a latitudes of $b=87^{\circ}~\&~ 80^{\circ}$ and serve to indicate the rate of this transition. These kernels are invariant under changes in longitude of the central pixel, the latitude being held fixed, as one may have expected.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Quantifying the non-locality of E \& B modes} \label{sec:radial_locality}
\subsection{The non-locality of the real space operators} \label{sec:radial_locality}
%
\begin{figure}[t]
\centering
\includegraphics[width=0.8\columnwidth]{beta_kernel.pdf}
\caption{The figure depicts the radial part of the convolution kernels. These radial function have been evaluated with the band limit fixed at $\ell \in [2,24]$. The vertical dashed line marks the approximate Healpix pixel size of a NSIDE=8, which is the lowest resolution that allows access to $\ell_{\rm max}=24$.}
\label{fig:beta_kernel}
\end{figure}
%
In the previous section we presented a quantitative understanding of the  azimuthal dependence of the real space kernels.  Here we probe the radial part that determines the non-locality of the operators and encodes all their multipole dependencies. For illustration the radial kernels ${_{\mm}f}, {_{\md}f},{_{\mi}f}$ are evaluated using the respective multipole sums given in \eq{eq:rad_ker_queb} and \eq{eq:f2_rad_ker} in the band limit $\ell \in [2,192]$ \revisit{not match figure?} and the resultant functions are depicted in \fig{fig:beta_kernel}. 
%
\begin{figure}[t]
\centering
\subfigure{\includegraphics[width=0.96\columnwidth]{rad_ker_fn_of_ellmax.pdf}}
\subfigure{\includegraphics[width=0.48\columnwidth]{rad_ker_i_fn_of_ellmax.pdf}}
\subfigure{\includegraphics[width=0.48\columnwidth]{rad_ker_d_fn_of_ellmax.pdf}}
\caption{The top panel depicts the radial function  ${_{\mm}f}(\beta,\ell_{\rm min},\ell_{\rm max})$ while the bottom left and right panels show the radial functions ${}_{\mi}f(\beta,\ell_{\rm min},\ell_{\rm max}) ~\&~ {}_{\md}f(\beta,\ell_{\rm min},\ell_{\rm max})$ respectively, for fixed $\ell_{\rm min}=2$ and varying $\ell_{\rm max}$ as indicated by their legends. All the curves are normalized such that their maxima is set to unity. The horizontal solid black line marks the location where the amplitude of the respective kernels fall below 1\% of its maximum. The thin slanted dashed gray lines indicate a power law fit (by eye) to the envelope of the radial functions. The thick black short vertical dashed lines indicate the transition points as predicted by the empirically derived relation for the non-locality parameter $\beta_0=min(180,180 \frac{22}{\ell_{\rm max}})$.}
%While the envelopes for function ${_{\mm}f}(\beta)~\&~ {}_{\mi}f(\beta)$ are fit well by the power law $\propto \beta^{-1.5}$, the envelope for the function ${}_{\md}f(\beta)$ is seen to have a slightly steeper slope $\propto \beta^{-1.65}$.}
\label{fig:rad_ker_decay}
\end{figure}
%

${_{\mm}f}$ is the radial part of the kernel that translates the Stokes parameters $Q$/$U$ to scalars $E$/$B$ and vice versa.  It has a vanishing contribution as $\beta \rightarrow 0$ and $\beta \rightarrow \pi$. This can be understood as arising from the fact that the coordinate dependence of the Stokes parameters cannot be uniquely integrated out since the Euler angles $\alpha$, $\gamma$ are not uniquely defined at the locations $\beta=0,\pi$. This nature of ${_{\mm}f}$ is critical to ensure that the derived fields have the necessary spin properties.

\revisit{${_{\mi}f}$ is the radial part of the band limited delta function $\mi$ and expectedly has its maxima at $\beta=0$ and decays with increasing angular separation.  ${_{\md}f}$ has a vanishing value in the region where $\beta \rightarrow 0$ however it does not vanish at $\beta \rightarrow \pi$ as seen in \fig{fig:beta_kernel}. The main takeaway from these observations of the nature of the radial part of the kernel is that the real space operators which decompose the Stokes parameters into parts that correspond to E and B modes are necessarily non-local.}\\ \linebreak

\paragraph{Band limit dependence.} 
%It is clear from previous discussions that the scalar field $E$ \& $B$ constructed at a location depends on the Stokes field in the surrounding regions. 
We further quantify this non-locality by studying the radial extent of the kernels and its dependence on the maximum multipole accessible for analysis. To carry out this study we evaluate the radial functions for different values of $\ell_{\rm max}$, while keeping the lowest multipole fixed at $\ell_{\rm min}=2$.
